### 2022-06-16 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@6f3e9e1ff2f027a1487065eab454118d787ca67e 🚀

In fig.~\ref{fig:biogeo}, we show a projection of the phylogeographic
signal of bats (top) and viruses (bottom) in space; the distinct
groupings (represented by different colors symbolizing positions in the
subspace formed by the first two axes of the PCoA) are essentially
equivalent between the two groups, and can be coarsely delineated as
southeast Asia, Eurasia above a northing of 25, and Africa and south
America. These results suggest that, although the evolutionary
distinctiveness of the bat/betacoronavirus complex varies spatially, the
system shows an important degree of spatial consistency, with a reduced
number of bioregions. Available information describing the spillover of
zoonotic betacoronaviruses of bat origin where data was available before
and up through the COVID-19 pandemic puts spillover events of SARS-CoV-2
in Wuhan, China; SARS-CoV in \textbf{XXX} based on the presence of
closest known viruses circulating in nature, and a nearby location where
serological (antibody) evidence has indicated human exposure to
SARS-like viruses (Wang et al.~2018 Virologica Sinica); MERS-CoV in
\textbf{XXX} based on index cases available from a recently-published
compendium of cases (Ramshaw et al.~2019). For the latest event, most if
not all index cases are presumed to be camel-to-human transmission, and
the precise origin point (if it exists) of MERS-CoV in bats is
uncertain. Recent recombinant canine coronavirus spillover events in
Haiti (\textbf{ref}) and Europe (\textbf{ref}) are not relevant here, as
bats' involvement in these cycles of transmission have been supposed to
be non-existent.

@@

Based on the previous result, we extracted the yellow component of the
risk map (TP add methods), to provide a single measure of risk varying
between 0 and 1. This measure is presented in fig.~\ref{fig:risk}.
However, this maps the potential risk, which must be weighed by the
potential for contacts with humans. As a proxy for this measure, we used
the proportion of build/urban land from the EarthEnv dataset: this is a
reasonable proxy for the density of humans per unit area, which
increases the probability of pathogen spread more widely (Hazarie et al.
2021). Since human activity is required to amplify the frequency of
virus encounters and thus create areas of viral amplification, mapping
the potential risk against measures of land use is required to generate
a more actionable assessment of risk. This map is presented in
fig.~\ref{fig:compound}. Most of South America and Europe are at low
risk, as although densely populated, settlements tend to be in areas
with lower potential risk. However, this mapping reveals that South-East
Asia, the Indian subcontinent, and parts of sub-Saharan Africa, are at
high risk due to the overlap between built areas and bat communities
representing more opportunities for cross-species transmission of
betacoronaviruses.

@@

@@

@@

@@

@@

>>>>

In fig.~\ref{fig:biogeo}, we show a projection of the phylogeographic
signal of bats (top) and viruses (bottom) in space; the distinct
groupings (represented by different colors symbolizing positions in the
subspace formed by the first two axes of the PCoA) are essentially
equivalent between the two groups, and can be coarsely delineated as
southeast Asia, Eurasia above a northing of 25, and Africa and south
America. These results suggest that, although the evolutionary
distinctiveness of the bat/betacoronavirus complex varies spatially, the
system shows an important degree of spatial consistency, with a reduced
number of bioregions. Available information describing the spillover of
zoonotic betacoronaviruses of bat origin where data was available before
and up through the COVID-19 pandemic puts spillover events of SARS-CoV-2
in Wuhan, China; SARS-CoV in Guangdong, China based on the presence of
closest known viruses circulating in nature, and a nearby location where
serological (antibody) evidence has indicated human exposure to
SARS-like viruses (Wang et al. 2018); MERS-CoV in Saudi Arabia based on
index cases available from a recently-published compendium of cases
(Ramshaw et al. 2019). For the latest event, most if not all index cases
are presumed to be camel-to-human transmission, and the precise origin
point (if it exists) of MERS-CoV in bats is uncertain. Recent
recombinant canine coronavirus spillover events in Haiti (Lednicky et
al. 2021) and Europe (Vlasova et al. 2022) are not relevant here, as
bats' involvement in these cycles of transmission have been supposed to
be non-existent. These index cases fall within different phylogeographic
bioregions (fig.~\ref{fig:biogeo}), which further highlight the issue
that different host-virus sub-systems may lead to widespread emergence.

@@

Based on the previous result, we extracted the yellow component of the
risk map (TP add methods), to provide a single measure of risk varying
between 0 and 1. This measure is presented in fig.~\ref{fig:risk}.
However, this maps the potential risk, which must be weighed by the
potential for contacts with humans. As a proxy for this measure, we used
the proportion of build/urban land from the EarthEnv dataset: this is a
reasonable proxy for the density of humans per unit area, which
increases the probability of pathogen spread more widely (Hazarie et al.
2021). Since human activity is required to amplify the frequency of
virus encounters and thus create areas of viral amplification, mapping
the potential risk against measures of land use is required to generate
a more actionable assessment of risk. This map is presented in
fig.~\ref{fig:compound}. Most of South America and Europe are at low
risk, as although densely populated, settlements tend to be in areas
with lower potential risk. However, this mapping reveals that South-East
Asia, the Indian subcontinent, and parts of sub-Saharan Africa, are at
high risk due to the overlap between built areas and bat communities
representing more opportunities for cross-species transmission of
betacoronaviruses. In looking for the origins of SARS in China, Xu et
al. (2004) present serological evidence that strongest human-animal
contact results in higher risk of virus exposure, regardless of the
animal species, but that different types of contact had different
impacts. Ideally, finer-grained information about human activity (rather
than human presence through anthropisation) could allow to partition
this risk further.

@@

\leavevmode\hypertarget{ref-Lednicky2021IsoNov}{}%
Lednicky, J. A. et al. 2021. Isolation of a Novel Recombinant Canine
Coronavirus from a Visitor to Haiti: Further Evidence of Transmission of
Coronaviruses of Zoonotic Origin to Humans. - Clinical Infectious
Diseases: An Official Publication of the Infectious Diseases Society of
America: ciab924.

@@

\leavevmode\hypertarget{ref-Ramshaw2019DatGeo}{}%
Ramshaw, R. E. et al. 2019. A database of geopositioned Middle East
Respiratory Syndrome Coronavirus occurrences. - Scientific Data 6: 318.

@@

\leavevmode\hypertarget{ref-Vlasova2022AniAlp}{}%
Vlasova, A. N. et al. 2022. Animal alphacoronaviruses found in human
patients with acute respiratory illness in different countries. -
Emerging Microbes \& Infections 11: 699--702.

@@

\leavevmode\hypertarget{ref-Wang2018SerEvi}{}%
Wang, N. et al. 2018. Serological Evidence of Bat SARS-Related
Coronavirus Infection in Humans, China. - Virologica Sinica 33:
104--107.

@@

\leavevmode\hypertarget{ref-Xu2004EpiClu}{}%
Xu, R.-H. et al. 2004. Epidemiologic Clues to SARS Origin in China. -
Emerging Infectious Diseases 10: 1030--1037.
